### 2018-11-22 paper/ijcai/figures/comparison_intro.tex
ijcai template

\small
\lstset{
  numbers=left,
  stepnumber=1,    
  firstnumber=1,
  numberfirstline=true,
  frame=none
}
\begin{figure}[t]
    \centering
\resizebox{0.9\columnwidth}{!}{%
    \begin{tabular}{l}
        \hline
        E \  C \  O\\ 
        \hline
        \lstinputlisting[
        language=C,
        basicstyle=\ttfamily,
        tabsize=1,
        ]{figures/example1.java} \\
        \hline
    \end{tabular}
    }
\caption{\label{caption:code-snippet-intro}
Code excerpt from the original Critical Slicing
    paper~\cite{DeMillo:1996:CSS:229000.226310}. E: Executed, C:
    (Original) Critical Slicing, O: \orbs{}~\cite{Binkley:2014:OLP:2635868.2635893}.  Symbol ``|''
    indicates that a statement is in the slice.}
\end{figure}
\normalsize